---
title: "Research Notes 1 Escobar"
author: "Rafael Escobar"
date: "4/8/2021"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
```

# Hypothesis
BMI and health insurance type are related

# Outcome: BMXBMI
# Study variable: HealthInsurance
# Confounders: Age, Male, Race/Ethnicity

##improt libraries and data set
```{r}
library("stargazer")
library("sas7bdat")
library("haven")
```

## Read SAS data file
```{r}
nhanes =  read_spss("NHANES_RN1.sav")
write.table(nhanes, file = "nhanes.txt",  sep = "\t")
class(nhanes)
dim(nhanes)
names(nhanes)

```


## Create factor variable for Health Insurace and Race/Ethnicity

```{r}
Ins_levels <- attributes(nhanes$HealthInsurance)$labels
nhanes$HealthInsurance_f <- factor(nhanes$HealthInsurance,
    levels = Ins_levels, labels = names(Ins_levels))
table(nhanes$HealthInsurance_f)
Race_levels <- attributes(nhanes$RaceEthnicity1)$labels
nhanes$Race_f <- factor(nhanes$RaceEthnicity1, levels = Race_levels, labels = names(Race_levels))
```
## Complete cases

```{r}
cc <- complete.cases(
    nhanes[, c("BMXBMI", "Male", "agegroup", "HealthInsurance_f", "Race_f")]
)
nhanes_sub <- nhanes[cc, ]
head(nhanes_sub)
```

### Polytomous Categorical Variable HealthInsurance to generate Binary Categorical Variable
```{r}
nhanes_sub$HealthInsurance_Private  <-  ifelse(nhanes_sub$HealthInsurance == 1, 1, 0)
```

### Looking at BMI
```{r}
summary(nhanes$BMXBMI)
table(nhanes$HealthInsurance)
dim(nhanes$HealthInsurance_f)
hist(nhanes$BMXBMI, breaks = 25, col = "lightgray", xlab = "BMI (kg/m2)", main = "Distribution and Probability Plot for BMI")
boxplot(nhanes$BMXBMI)
qqnorm(nhanes$BMXBMI, ylab = "BMI (kg/m2)")
qqline(nhanes$BMXBMI, col= "steelblue", lwd= 2)

### Log Transformation of BMI to address skewness 
nhanes_sub$logBMI <- log(nhanes_sub$BMXBMI)
qqnorm(nhanes_sub$logBMI)
qqline(nhanes_sub$logBMI, col = "steelblue", lwd = 2)
hist(nhanes_sub$logBMI, breaks = 25, col = "lightgray", xlab = "LogBMI", main = "Distribution and Probablity Plot for logBMI")
summary(nhanes_sub$logBMI)
sd(nhanes_sub$logBMI)

### observation of normality 
plot(nhanes_sub$Exactage , nhanes_sub$BMXBMI, xlab = "Exact Age", ylab = "BMI (kg/m2)")
plot(nhanes_sub$BMXBMI ~ nhanes_sub$Exactage, xlab= "Exact Age", ylab= "BMI (kg/m2)")
plot(nhanes_sub$BMXBMI ~ nhanes_sub$agegroup,xlab= "Age Groups", ylab= "BMI (kg/m2)")
plot(nhanes_sub$logBMI~ nhanes_sub$Exactage,xlab= "Exact Age", ylab= "LogBMI")
plot(nhanes_sub$logBMI ~ nhanes_sub$agegroup, xlab= "Age Groups", ylab= "LogBMI")
hist(nhanes_sub$Exactage, breaks = 25, col = "lightgray", xlab = "Exact Age", main = "Histogram of Exact Age")
qqnorm(nhanes_sub$Exactage)
hist(nhanes_sub$agegroup, breaks = 25, col = "lightgray", xlab = "Age Groups", main = "Histogram of Age Groups")
boxplot(nhanes_sub$logBMI ~ nhanes_sub$agegroup, xlab= "Age Groups", ylab = "logBMI", main = "Distribution of logBMI by 'agegroup'")
```

#### Unadjusted Model

```{r}
unadjusted <- lm( logBMI ~ HealthInsurance_Private, data = nhanes_sub)
stargazer(unadjusted, title="Unadjusted",
   type="text", style="default", header=FALSE)
```
### Adjusted Model

```{r}
adjusted <- lm( logBMI ~ HealthInsurance_Private + RIAGENDR +  factor(agegroup) + Race_f, data= nhanes_sub)
stargazer(adjusted, title="Adjusted",
   type="text", style="default", header=TRUE, object.names = TRUE, covariate.labels = c("PrivateInsurance", "Male*", "20-24*", "25-29", "30-34", "35-39", "40-44", "45-49", "50-54", "Non-Hispanic White*", "Non-Hispanic Black", "Hispanic")
                               )
```
### F-test between two models
```{r}
anova(adjusted, unadjusted)
```


##### Obtaining the fitted values and residual values:Unadjusted

```{r}
res <- residuals(unadjusted)
yhat <- fitted(unadjusted)
```
### Plotting fitted vs residuals

```{r}
hist(res,breaks = 25, col = "lightgray", xlab = "Residuals", main = "Distribution and Probability Plot for Residuals")
plot(yhat, res, xlab = "Predicted Y", ylab = "Resuduals")
```

##### Normality of Residuals

### Quantile-Quantile Plot

```{r}
qqnorm(res)
qqline(res, col = "steelblue", lwd = 2)
```

###### Obtaining the fitted values and residual values: Adjusted
```{r}
res2 <- residuals(adjusted)
yhat2 <- fitted(adjusted)
```
### Plotting fitted vs residuals

```{r}
hist(res2,breaks = 25, col = "lightgray", xlab = "Residuals", main = "Distribution and Probability Plot for Residuals")
plot(yhat2, res2, xlab = "Predicted Y", ylab = "Resuduals")
```

## Normality of Residuals

To check for normality we can plot our distribution values (normal distribution)
versus our residuals in a Q-Q plot.

### Quantile-Quantile Plot

```{r}
qqnorm(res2)
qqline(res2, col = "steelblue", lwd = 2)
```

